---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# resquin

<!-- badges: start -->

[![CRAN
status](https://www.r-pkg.org/badges/version/resquin)](https://CRAN.R-project.org/package=resquin)

<!-- badges: end -->

## About

**`resquin` is currently in development and should only be used with
caution. This message will be removed once the package is released on CRAN.**

`resquin` (**res**ponse **qu**ality **in**dicators) provides functions
to calculate survey data quality indicators to help identifying
low-quality responses [@bhaktha; @curran2016; @vanvaerenbergh2013].
`resp_styles()` and `resp_distributions()` provide response quality indicators geared towards multi-item
scales or matrix questions. Both multi-item scales and matrix questions
present survey respondents with multiple questions which have the same response format,
meaning the same number and labeling of response options.

At the moment, `resquin` provides two functions:

-   `resp_styles()` - Calculates response style indicators (e.g. extreme
    response style or middle response style).
-   `resp_distributions()` - Calculates response distribution indicators
    (e.g. intra-individual mean and standard deviation over a set of
    survey questions).

Two more functions are planned:

-   `resp_patterns` - Calculates response pattern indicators (e.g.
    straightlining)
-   `resp_times` - Calculates response time indicators (e.g. median item
    response time)
    
For information on how to use `resquin` see the vignettes [Getting started with resquin](https://matroth.github.io/resquin/articles/getting_started_with_resquin.html) and [resquin in practice](https://matroth.github.io/resquin/articles/resquin_in_practice.html).

## Installation

`resquin` is available via CRAN and github. To install `resquin` from CRAN or 
github, you can use one of the following commands:

```{r eval=FALSE}
# Install resquin via CRAN
install.packages("resquin")

# Install development version of resquin with devtools
devtools::install_github("https://github.com/MatRoth/resquin")

# Install development version of resquin with pak
pak::pak("https://github.com/MatRoth/resquin")
```

## Getting started

To use `resquin`, supply a data frame containing survey responses in wide format 
to either `resp_styles()` or `resp_distributions()`.

```{r}
# load resquin
library(resquin)

# A test data set with three items and ten respondents
testdata <- data.frame(
  var_a = c(1,4,3,5,3,2,3,1,3,NA),
  var_b = c(2,5,2,3,4,1,NA,2,NA,NA),
  var_c = c(1,2,3,NA,3,4,4,5,NA,NA))

testdata

# Calculate response style indicators per respondent
resp_styles(x = testdata,
            scale_min = 1,
            scale_max = 5) |> # Specify scale minimum and maximum
  round(2)

# Calculate response distribution indicators per respondent
resp_distributions(x = testdata) |>
  round(2)
```

For a more information on how to use `resquin` see the vignettes [Getting started with resquin](https://matroth.github.io/resquin/articles/getting_started_with_resquin.html) and [resquin in practice](https://matroth.github.io/resquin/articles/resquin_in_practice.html).

# References
