---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# resquin

<!-- badges: start -->

[![CRAN
status](https://www.r-pkg.org/badges/version/resquin)](https://CRAN.R-project.org/package=resquin)
[![R-CMD-check](https://github.com/MatRoth/resquin/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/MatRoth/resquin/actions/workflows/R-CMD-check.yaml)
[![Codecov test
coverage](https://codecov.io/gh/MatRoth/resquin/graph/badge.svg)](https://app.codecov.io/gh/MatRoth/resquin)

<!-- badges: end -->

## About

`resquin` (**res**ponse **qu**ality **in**dicators) provides functions
to calculate survey data quality indicators to help identifying
low-quality responses [@bhaktha; @curran2016; @vanvaerenbergh2013].
`resp_styles()`, `resp_distributions()`, `resp_nondifferentiation()` and
`resp_patterns()` provide response quality indicators geared towards
multi-item scales or matrix questions. Both multi-item scales and matrix
questions present survey respondents with multiple questions which have
the same response format, meaning the same number and labeling of
response options.

At the moment, `resquin` provides four functions:

-   `resp_styles()` - Calculates response style indicators (e.g. extreme
    response style or middle response style).
-   `resp_distributions()` - Calculates response distribution indicators
    (e.g. intra-individual mean and standard deviation over a set of
    survey questions).
-   `resp_nondifferentiation()` - Calculates response nondifferentiation
    indicators. Nondifferentiation indicators primarily measure
    straightlining. The indicators differ in how straightlining is
    operationalized.
-   `resp_patterns()` - Calculates response pattern indicators (e.g.
    long string analysis).

For information on how to use `resquin` see the vignettes [Getting
started with
resquin](https://matroth.github.io/resquin/articles/getting_started_with_resquin.html)
and [resquin in
practice](https://matroth.github.io/resquin/articles/resquin_in_practice.html).

`resquin` is still under active development. Please use github
[issues](https://github.com/MatRoth/resquin/issues) to file questions
and bug reports or send them directly to
[matthias.roth\@gesis.org](mailto:matthias.roth@gesis.org){.email}. We
are happy to receive feedback!

## Installation

`resquin` is available via CRAN and github. To install `resquin` from
CRAN or github, you can use one of the following commands:

```{r eval=FALSE}
# Install resquin via CRAN
install.packages("resquin")

# Install development version of resquin with devtools
devtools::install_github("https://github.com/MatRoth/resquin")

# Install development version of resquin with pak
pak::pak("https://github.com/MatRoth/resquin")
```

## Getting started

To use `resquin`, supply a data frame containing survey responses in
wide format to either `resp_styles()`, `resp_distributions()`,
`resp_nondifferentiation()`, `resp_patterns()`.

```{r}
# load resquin
library(resquin)

# A test data set with three items and ten respondents
testdata <- data.frame(
  var_a = c(1,4,3,5,3,2,3,1,3,NA),
  var_b = c(2,5,2,3,4,1,NA,2,NA,NA),
  var_c = c(1,2,3,NA,3,4,4,5,NA,NA))

testdata

# Calculate response style indicators per respondent
resp_styles(x = testdata,
            scale_min = 1,
            scale_max = 5) |> # Specify scale minimum and maximum
  round(2)

# Calculate response distribution indicators per respondent
resp_distributions(x = testdata) |>
  round(2)

# Calculate response nondifferentiation indicator per respondent
resp_nondifferentiation(x = testdata) |> 
  round(2)

# Calculate response pattern indicators
resp_patterns(x = testdata) |> 
  round(2)
```

For a more information on how to use `resquin` see the vignettes
[Getting started with
resquin](https://matroth.github.io/resquin/articles/getting_started_with_resquin.html)
and [resquin in
practice](https://matroth.github.io/resquin/articles/resquin_in_practice.html).

Researchers have developed a large number of response quality
indicators. `resquin` only covers a few of them. The following list
contains packages which provide alternative implementations of the
indicators used in `resquin` or provide other indicators not used in
`resquin`:

-   Implementation of [@kim_straightlining_2019] measures:
    [straightliner](https://github.com/mweylandt/straightliner/tree/main)

-   Implementation of indices from the careless responding literature:
    [careless](https://cran.r-project.org/package=careless)

-   Autocorrelation and pattern indices:
    [responsePatterns](https://cran.r-project.org/package=responsePatterns)

Below are packages which use IRT to model and estimate response styles
and individual response quality:

-   [PCMRS](https://cran.r-project.org/package=PCMRS)

-   [FitResponseStyles](https://github.com/mirka-henninger/FitResponseStyles)

-   [Model Multivariate Ordinal Responses Including Response
    Styles](https://cran.r-project.org/package=MultOrdRS)

-   [Person fit indices
    (Perfit)](https://cran.r-project.org/package=PerFit)

# References
