---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# resquin

<!-- badges: start -->
<!-- badges: end -->

## About

`resquin` (**re**sponse **qu**ality **in**dicators) provides functions to
calculate common survey data quality indicators, such as response style indicators. 

At the moment, `resquin` provides two functions: 

* `resp_distributions()` - Calculates response quality indicators (e.g. within person 
mean and standard deviation over a set of survey questions).
* `resp_styles()` - Calculates response style indicators (e.g. extreme response style
or middle response style).

Two more functions are planned: 

* `resp_patterns` - Calculates response pattern indicators (e.g. straightlining)
* `resp_times` - Calculates response time indicators (e.g. median item response time)

## Installation

`resquin` is available via github.
To install `resquin` from github, you can use one of the following commands:

```{r eval=FALSE}   
# Installing resquin with devtools
devtools::install_github("https://github.com/MatRoth/resquin")

# Installing resquin with pak
pak::pak("https://github.com/MatRoth/resquin")
``` 

## Getting started

Functions in `resquin` calculate response quality indicators for survey data stored in a data frame or 
tibble. The functions assume that the input data frame is structured in the following way:

* The data frame is in wide format, meaning each row represents one respondent, each
column represents one variable.
* All variables have the same number of response options.
* The variables are in same the order as the questions respondents saw while taking the survey.
* Reverse keyed variables are in their original form. No items were recoded.
* All responses have integer values.
* Missing values are set to `NA`.

### Example dataset of survey responses

Consider the following (fake) data set of survey responses. 

```{r}
# A small test data set with three items and ten respondents
testdata <- data.frame(
  var_a = c(1,1,4,3,5,3,2,3,1,3,NA),
  var_b = c(2,2,5,2,3,4,1,NA,2,NA,NA),
  var_c = c(1,2,2,3,NA,3,4,4,5,NA,NA))

testdata
```
The data set contains responses to three survey questions (var_a,var_b and var_c) from ten respondents.
All three survey question allow responses on a scale from 1 to 5. Some respondents have
missing values, which are set to `NA`.

Lets use this data set to calculate response quality indicators. 

### `resp_styles()`: Response style indicators

Response styles capture systematic shifts in respondents response behavior.
For example, respondents with an extreme response style may only choose the lowest and highest categories (in our example 1 and 5)
while mid-point responder only choose the midpoint of a scale (in our example 3). 

To calculate response styles we can use the `resp_styles()` function.
First, we need to specify our data argument `x`. Then, we need to specify the minimum and maximum of the scales
used in our questionnaire (`scale_min` and `scale_max` respectively). Remember that all questions included must have the same number of response options.
We will discuss the arguments `min_valid_responses` and `normalize` later.

```{r}
library(resquin)
# Calculating response style indicators for all respondents with no missing values
results_response_styles <- resp_styles(
  x = testdata,
  scale_min = 1,
  scale_max = 5,
  min_valid_responses = 1, # Excludes respondents with less than 100% valid responses
  normalize = T)  # Presents results in percent of all responses

round(results_response_styles,2)
```

The resulting data frame contains five columns corresponding to the middle response style (MRS),
acquiescence response style (ARS), disaquiescence response style (DRS), extreme response style (ERS),
and non-extreme response style (NERS) - you can learn more about the response styles
in the help file of the function using `?resp_styles`.

Each respondent receives one value for each indicator, given that they can be calculated. Because 
`normalize` is set to `TRUE`the values are expressed as the share of responses of a respondent 
that can be attributed to a response style. For example, respondent one has an ERS value of 0.67
meaning that two out of three responses can be identified as extreme responses.
On the other hand, respondent one does not have any 
mid-point response, leading to a value of 0 in the MRS column. 

Instead of calculating proportions, we can extract the counts of responses that can
be attributed to a response option by setting `normalize` to `FALSE`.

Finally, we can decide to include or exclude respondents from receiving response
style values by setting `min_valid_responses`, which can take values from 0 to 1. `min_valid_responses` sets the
share of valid responses (i.e. non-missing responses) a respondent must have to receive
response style values. A value of 0 indicates that response style values should be
calculated for all respondents, regardless of whether or not they have missing values. A value 
of 1 indicates that response styles should only be calculated if respondents have
valid responses on all variables. Values between 0 and 1 indicate the share of
responses that need to be valid to be included in the response style calculations.

### `resp_distributions()`: Within respondent response distribution indices 

`resp_distributions()` calculates indices which reflect the location and variability of responses within a respondent.

```{r}
# Calulating response distribution indicators for all respondents with no missing values
resp_distributions(testdata,min_valid_responses = 1) |>
  round(2) # Round to second decimal place
```
The resulting data frame contains eight columsn:
  * n_valid: the number of valid responses
  * n_na: the number of missing responses
  * prop_na: the proportion of missing responses of all responses
  * ips_mean: the ipsatized (within respondent) mean over all responses
  * ips_median: the ipsatized (within respondent) median over all responses
  * ips_sd: the ipsatized (within respondent) standard deviation over all responses
  * mahal: the mahalanobis distance of the respondent across all responses




