---
title: "Flagging respondents"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Flagging respondents}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Using `flag_resp()` to create and compare flagging strategies
One use-case for response quality indicators is to use them to flag responses
which potentially are of low quality. `resquin` provides the function `flag_resp()`
to create a data frame of booleans (`T` and `F`) according to user-defined cut-off values on 
response quality indicators. If a respondent receives a `T` value, they are flagged
as suspicious. If they receive `F` value, they are deemed unsuspicious.

The strength of `flag_resp()` lies in its ability to quickly create and compare
multiple flagging strategies, as the following example illustrates:

Suppose we use data on response styles to decide whether respondents are low-quality
responders on the 15 item `nep` scale. We can use `resp_styles()` to calculate response style indices
per respondent.


```{r setup}
library(resquin)
nep_resp_styles <- resp_styles(
  x = nep,
  scale_min = 1, # minimum response option
  scale_max = 5, # maximum response option
  min_valid_responses = 1) # default, excludes respondents with any missing value

summary(nep_resp_styles)
```

In the first example, we will consider the acquiescence response style (ARS).
ARS represents the tendency of respondents to agree to questions regardless of
their content. Since the `nep` scale includes positively and negatively keyed items, 
we can expect that higher ARS values indeed correspond to this behavior: Respondents
who are more concerned about nature should choose higher response options on the positively
keyed items and more negative responses on the negatively keyed items. Just choosing 
all high response options presents a substantively inconsistent response behavior,
potentially caused by acquiescence.

A first idea could be to flag respondents which have more than 80% responses in the
ARS category. 

```{r}
first_flagging <- flag_resp(nep_resp_styles,
                            ARS > 0.8)

summary(first_flagging)
```
We can see that 33 respondents are flagged as suspicious, as their ARS score is above 0.8. In a second step, we might also be interested in those respondents who choose the middle response option a lot. The mid-point response style (MRS) could be a sign of speeding or satisficing, as respondents do not make a choice in either direction of the response scale. We keep the ARS flagging strategy in place to compare it to the new one.

```{r}
second_flagging <- flag_resp(nep_resp_styles,
                             ARS > 0.8,
                             MRS > 0.8)
summary(second_flagging)
```
We can see that only three respondents have a MRS value above 0.8. Unsurprisingly, no respondent who has a ARS > 0.8 also has a MRS > 0.8.

