---
title: "Getting started with resquin"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{resquin tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## An introduction to `resquin`

This short tutorial describe the functions in `resquin` and how you can use them
on a technical level. For a more substantive introduction see the (forthcoming) 
article [Using resquin in practice](https://matroth.github.io/resquin/articles/resquin_in_practice.html).

Functions in `resquin` calculate response quality indicators for survey
data stored in a data frame or tibble. The functions assume that the
input data frame is structured in the following way:

-   The data frame is in wide format, meaning each row represents one
    respondent, each column represents one variable.
-   All variables have the same number of response options.
-   The variables are in same the order as the questions respondents saw
    while taking the survey.
-   All responses have integer values.
-   Missing values are set to `NA`.
-   (For `resp_styles()`) Reverse keyed variables are in their original form. No items were
    recoded.

### Example dataset of survey responses

Consider the following (fake) data set of survey responses.

```{r}
# A test data set with three items and ten respondents
testdata <- data.frame(
  var_a = c(1,4,3,5,3,2,3,1,3,NA),
  var_b = c(2,5,2,3,4,1,NA,2,NA,NA),
  var_c = c(1,2,3,NA,3,4,4,5,NA,NA))

testdata
```

The data set contains responses to three survey questions (var_a,var_b
and var_c) from ten respondents. All three survey question allow
responses on a scale from 1 to 5. Some respondents have missing values,
which are set to `NA`.

Lets use this data set to calculate response quality indicators.

### `resp_styles()`: Response style indicators

Response styles capture systematic shifts in respondents response
behavior. For example, respondents with an extreme response style may
only choose the lowest and highest categories (in our example 1 and 5)
while mid-point responder only choose the midpoint of a scale (in our
example 3).

To calculate response styles we can use the `resp_styles()` function.
First, we need to specify our data argument `x`. Then, we need to
specify the minimum and maximum of the scales used in our questionnaire
(`scale_min` and `scale_max` respectively). Remember that all questions
included must have the same number of response options. We will discuss
the arguments `min_valid_responses` and `normalize` later.

```{r}
library(resquin)
# Calculating response style indicators for all respondents with no missing values
results_response_styles <- resp_styles(
  x = testdata,
  scale_min = 1,
  scale_max = 5,
  min_valid_responses = 1, # Excludes respondents with less than 100% valid responses
  normalize = T)  # Presents results in percent of all responses

round(results_response_styles,2)
```

The resulting data frame contains five columns corresponding to the
middle response style (MRS), acquiescence response style (ARS),
disaquiescence response style (DRS), extreme response style (ERS), and
non-extreme response style (NERS) - you can learn more about the
response styles in the help file of the function using `?resp_styles`.

Each respondent receives one value for each indicator, given that they
can be calculated. Because `normalize` is set to `TRUE`the values are
expressed as the share of responses of a respondent that can be
attributed to a response style. For example, respondent one has an ERS
value of 0.67 meaning that two out of three responses can be identified
as extreme responses. On the other hand, respondent one does not have
any mid-point response, leading to a value of 0 in the MRS column.

Instead of calculating proportions, we can extract the counts of
responses that can be attributed to a response option by setting
`normalize` to `FALSE`.

Finally, we can decide to include or exclude respondents from receiving
response style values by setting `min_valid_responses`, which can take
values from 0 to 1. `min_valid_responses` sets the share of valid
responses (i.e. non-missing responses) a respondent must have to receive
response style values. A value of 0 indicates that response style values
should be calculated for all respondents, regardless of whether or not
they have missing values. A value of 1 indicates that response styles
should only be calculated for respondents who have valid responses on
all variables. Values between 0 and 1 indicate the share of responses
that need to be valid to be included in the response style calculations.

### `resp_distributions()`: Intra-individual response distribution indicators

`resp_distributions()` calculates indicators which reflect the location
and variability of responses within a respondent. `resp_distributions()`
works similar to `resp_styles()`: We need to specify the data argument
and we can include or exclude respondents from the calculations based on
amount of missing data they exhibit (for an explanation see paragraph
above).

```{r}
# Calulating response distribution indicators for all respondents with no missing values
results_resp_distributions <- resp_distributions(
  x = testdata,
  min_valid_responses = 1) # Excludes respondents with less than 100% valid responses

round(results_resp_distributions,2)
```

The resulting data frame contains eight columns:

-   n_na: number of intra-individual missing answers

-   prop_na: proportion of intra-individual missing responses

-   ii_mean: intra-individual mean

-   ii_median: intra-individual median

-   ii_sd: intra-individual standard deviation

-   mahal: Mahalanobis distance per respondent.

You can learn more about the response distribution indicators using `?resp_distributions`

### `resp_nondifferentiation()`: Response nondifferentiation indicators
`resp_nondifferentiation` calculates indicators primarily measuring what is known
as straightlining. However, different indicators measure different aspects of straightlining.
While *Simple Nondifferentiation* just measures whether respondents used the same response
option for all questions, other indicators such as the *Mean Root of Pairs Method* or 
the *Scale Point Variation Method* provide continuous values of nondifferentiation,
measuring how varied the choice of response options for each respondent is.

`resp_nondifferentiation()` works just like `resp_distributions()` and `resp_styles()`.
```{r}
results_resp_nondifferentiation <- resp_nondifferentiation(
  x = testdata,
  min_valid_responses = 1) # Excludes respondents with less than 100% valid responses

round(results_resp_nondifferentiation,2)
```

The resulting dataframe contains four columns containing indicators described
by [@kim_straightlining_2019]:

-   simple_nondifferentiation: Respondents are assigned 1 or 0 depending on 
    whether all responses have the same value (1) or not (0).
-   mean_root_pairs: Mean of the root of the absolute differences between
    all pairs in a multi-item scale or matrix questions. It ranges from 0
    (least straightlining) to 1 (most straightlining). 
-   max_identical_rating: Proportion of the most commonly selected
    response option among all responses in a multi-item scale or matrix questions.
    It ranges from 0 (least straightlining) to 1 (most straightlining).
-   scale_point_variation: The measure becomes larger if respondents use
    more scales points in a multi-item scale or matrix questions.

You can learn more about the the response nondifferentiation indicators using 
`?resp_nondifferentiation`.

### `resp_patterns()`: Response pattern indicators
Response patterns describe suspicious response patterns arising from careless or insufficient effort responding. Careless or insufficient effort respondents might choose the same answer option repeatedly (i.e. creating a long string of equal responses) or use pattern, such as zig-zaging, to complete a survey.

`resp_patterns()` provides three columns which contain longstring analysis indicators for each respondent. These three columns are always returned. There are two more optional columns () for the analysis of other repeating response patterns: defined_patterns and arbitrary_patterns:

To obtain the three longstring analysis columns simply call `resp_patterns()`:
```{r}
results_resp_patterns <- resp_patterns(testdata)
round(results_resp_patterns,2)
```

To obtain counts of defined response patterns per respondents use the `defined_patterns`
argument. You can either supply a single vector representing a response pattern, 
or a list of vectors representing multiple response patterns. 

```{r}
# Single defined pattern
defined_patterns_single <- resp_patterns(
  x = testdata,
  defined_patterns = c(1,2,3)
)
defined_patterns_single

# Multiple defined patterns
defined_patterns_multiple <- resp_patterns(
  x = testdata,
  defined_patterns = list( # wrap multiple pattern vectors into a list
    c(1,2,3),
    c(2,3,4),
    c(4,3,2),
    c(3,2,1)
  )
)

defined_patterns_multiple

# defined patterns are returned as a single list column.
# One way to make the data accessible is by using tidyr::unnest_wider()
# This way, each column represents the count of one defined pattern

defined_patterns_multiple |> 
  tidyr::unnest_wider(defined_patterns)
```

Finally, you can request the detection of arbitrary response patterns with counts
larger than one and pattern lengths longer than one using the `arbitrary_patterns`
argument

```{r}
arbitrary_patterns_length_2 <- resp_patterns(
  x = testdata,
  arbitrary_patterns = 2
)

arbitrary_patterns_length_2

# You can also request the detection of patterns of multiple 
# lengths, in this case 2 and 3
arbitrary_patterns_length_2_3 <- resp_patterns(
  x = testdata,
  arbitrary_patterns = c(2,3)
)
arbitrary_patterns_length_2_3
```



### Using the `id` argument: Uniquely identify respondents

In its default setting, all `resquin` functions provide an integer `id` column
running from 1 to the number of respondents in the data set supplied in `x`. You can turn of the
creation of the `id` column by setting `id = False` or supply a vector of unique
ids (character or numeric). The latter option is useful if a data set provides its
own vector of ids.

Here are examples of the three options:
```{r}
# Default. Integer ids.
resp_distributions(testdata)

# No id column
resp_distributions(testdata,id = F)

# Custom id vectors
custom_ids <- letters[1:nrow(testdata)]
resp_distributions(testdata,id = custom_ids)
```

### Missing data as a threat to the validity of response quality indicators

The argument `min_valid_responses` can be used to control how many 
responses a respondent must have for a response quality indicator to be computed.
Per default, `resquin` requires respondents to have valid responses for all questions
asked `(min_valid_responses = 1)`. In other words, per default `resquin` requires
respondents to have no missing values. 

Of course, missing values are a common occurrence in survey data, thus it can make
sense to reduce the required share of valid responses in `min_valid_responses` from 1 to a
lower one. This leads to the calculation of response quality indicator values 
for more respondents. For example, reducing `min_valid_responses` to 0.5 would mean
that respondents only have to have valid (i.e. non-missing) responses on 50% of
the questions asked. 

However, allowing `NA` values (i.e. missing values) can change the interpretation
and validity of response quality indicators.
For example, a respondent with ten identical responses on a ten item scale
(i.e. ten time response option "3") has more evidence for straightlining then a respondent with two identical responses
which are separated by eight `NA` values. Both would receive a 1 on the 
`simple_nondifferentiation` indicator of `resp_nondifferentiation()` or 1 (meaning
absolute presence of) a Middle Response Style if the response scale had five response options.
But should the respondents be considered to be identical in their extent of
straightlining or Middle Response Style? This is up to the interpretation of the 
analyst. 

Although an increasing number of missing values per respondent can pose validity
threats to all response quality indicators, some indicators are more affected 
by `NA` values than others. Indicators calculated with `resp_distributions()` are
not affected as much by missing data, because the `resp_distributions()` indicators
average over all response values.
For the following response indicators missing values are more of a threat to validity
because they conceptually rely more on the position of responses on the response
scale:

* `resp_patterns()`:
    + `mean_string_length` - Response strings with identical values can be broken by `NA`.
    + `longest_string_length` - Response strings with identical values can be broken by `NA`.

* `resp_nondifferentiation()`:
    + `simple_nondifferentiation` - Straightlining with identical values can be broken by `NA`.
  
* `resp_styles()`:
    + All response styles - Response strings with identical values can be broken by `NA`.

In any case, it helps to report how missing values were handled in an analysis to 
increase confidence in and replicability of the results. 

### References


